# Methods

<!-- The Methods should include detailed text describing any steps or procedures used in producing the data, including full descriptions of the experimental design, data acquisition and any computational processing. See [the detailed section in our submission guidelines](https://www.nature.com/sdata/publish/submission-guidelines#sec-5) for advice on writing a transparent and reproducible methods section. Related methods should be grouped under corresponding subheadings where possible, and methods should be described in enough detail to allow other researchers to interpret and repeat, if required, the full study. Specific data outputs should be explicitly referenced via data citation (see Data Records and Citing Data, below). -->

<!-- Authors should cite previous descriptions of the methods under use, but ideally the method descriptions should be complete enough for others to understand and reproduce the methods and processing steps without referring to associated publications. There is no limit to the length of the Methods section. -->

This section provides a comprehensive description of the steps and procedures used to create the Nominal catch dataset from the five different datasets, each representing one tRFMO (Regional Fisheries Management Organization).

# Data origin

## Gathering of the data by tRFMOs:

Each tRFMOs has different exchange formats to gather data from fishboats/countries or fisheries.
to be continued

## Gathering for Nominal catch dataset:

### Data call

The 5 dataset used for creating the Nominal catch dataset is gathered through a systematic datacall process conducted annually by the FIRMS (Fisheries Resources Monitoring System) steering committee. This datacall ensures the collection of the most up-to-date and comprehensive information on tuna catch from the 5 tRFMOs (CCSBT, IATTC, ICCAT, IOTC, WCPFC). The datacall typically takes place around June each year, allowing for the timely availability of the data for analysis.

In addition to the annual datacall put in place by the FIRMS steering committee, the dataset can also be updated periodically from the official websites provided by the tRFMOs.

### The CWP Data Exchange Format

For effective amalgamation of multiple datasets, a unified data exchange format has been crafted. Central to this format is the inclusion of:

-   Digital Format Standardization:

The Comma-Separated-Value (CSV) format, adhering to the RFC 4180 standard, has been endorsed.

-   File Encoding Protocol:

Whether in CSV or JSON format, encoding is standardized to UTF-8.

-   Data Dictionary:

A robust data dictionary has been established to guide the structure and semantics of the fisheries data exchange format. This repository, containing terms for column headers and their definitive explanations, is accessible at FDI Terms on GitHub. <https://github.com/fdiwg/fdi-terms> Notably, a specific fisheries term isn't rigidly bound to a single reference code list but varies based on its application context.

In context-specific scenarios, like the Global Tuna Atlas, associations with particular code lists : <https://github.com/fdiwg/fdi-mappings/tree/main/global/firms/gta>

Lastly, while two data structure definitions are being created, the Global Tuna Atlas has adopted the Generic (measurement-independent) data structure for its operations. \ref table

```{r results='asis'}
CWP_format <- read_delim("inputs/CWP_format.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
qflextable(CWP_format)
```

# Creation of the Global Dataset

## Harmonization

For now the CWP Data Exchange Format has not been adopted by all the tRFMOs yet. Then a pre processing part of Digital Format Standardization is ran before the workflow. These step is to deaseppear in future release. reftable\

```{r  results='asis'}
require(readr)
datasets_links_processing <- read_delim(here("inputs/Workflow/datasets_links_processing.csv"), 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
require(flextable)
qflextable(datasets_links_processing)
```

## Final output

Upon completing the CWP standardization, the global Nominal catch dataset is constructed in R. It takes in input the 5 datasets provided by tRFMOS and process 3 treatments.

1.  Binding of Datasets for each Ocean Region:

All the individual tRFMO datasets, representing different ocean regions, are combined through a data-binding process.

2.  Mapping the data

If the data is not mapped yet, the mapping is done on the basis of the codelist presented sooner. The recap of the mapping can be found in table \ref table

3.  Filtering on Species under Mandate by the tRFMOs:

This step involves filtering the data to retain only the information pertaining to species under the mandate of the tRFMOs (Regional Fisheries Management Organizations). The retained species are selected to align with the specific conservation and management goals of each tRFMO.

```{r}

file_url <- "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/cross-term/codelist_mapping_source_authority_species.csv"
download_link <- paste0("[Download CSV](<", file_url, ">)")

cl_species_level0 <- readr::read_csv(file_url)


```

```{r}

diffstratasnumber <- read.csv("outputs/filtering_species/figures/Diffstratas/Comparison.of.number.of.stratas.between.the.two.datasets.csv")

tidy <- diffstratasnumber[c(1, 3, 4, 5, 6),] %>% rename(" " = X, "Before filtering" = "Before.filtering", "After filtering" = "After.filtering")

qflextable2(tidy, columns_to_color = "Difference")

```

<!-- To be noted, one of the main tuna i.e. Southern Bluefin Tuna, for which CCSBT is considered the only authoritative source of information for the species is impacted by 4.3% of the value by this treatment. -->

To enhance the comprehensibility and usability of the dataset, additional temporal and grouping information is enriched in the data, categorizing species and fishing gear types using mapping of the Data Dictionary.

The code used to perform this workflow is available on: [https://raw.githubusercontent.com/eblondel/geoflow-tunaatlas/master/tunaatlas_scripts/generation/create_global_tuna_atlas_nominal_catch.R](#0){.uri}

## Geoflow workflow

The workflow of creation and the dissemination of the Nominal catch is using the geoflow package <https://github.com/r-geoflow/geoflow>.
