# Methods


<!-- The Methods should include detailed text describing any steps or procedures used in producing the data, including full descriptions of the experimental design, data acquisition and any computational processing. See [the detailed section in our submission guidelines](https://www.nature.com/sdata/publish/submission-guidelines#sec-5) for advice on writing a transparent and reproducible methods section. Related methods should be grouped under corresponding subheadings where possible, and methods should be described in enough detail to allow other researchers to interpret and repeat, if required, the full study. Specific data outputs should be explicitly referenced via data citation (see Data Records and Citing Data, below). -->

<!-- Authors should cite previous descriptions of the methods under use, but ideally the method descriptions should be complete enough for others to understand and reproduce the methods and processing steps without referring to associated publications. There is no limit to the length of the Methods section. -->

This section provides a comprehensive description of the steps and procedures used to create the Nominal catch dataset from the five different datasets, each representing one tRFMO (Regional Fisheries Management Organization).

# Data origin

## Gathering of the data by tRFMOs:

Each tRFMOs has different exchange formats to gather data from fishboats/countries or fisheries.
to be continued

## Gathering for Nominal catch dataset:

### Data call

The 5 dataset used for creating the Nominal catch dataset is gathered through a systematic datacall process conducted annually by the FIRMS (Fisheries Resources Monitoring System) steering committee. This datacall ensures the collection of the most up-to-date and comprehensive information on tuna catch from the 5 tRFMOs (CCSBT, IATTC, ICCAT, IOTC, WCPFC). The datacall typically takes place around June each year, allowing for the timely availability of the data for analysis.

In addition to the annual datacall put in place by the FIRMS steering committee, the dataset can also be updated periodically from the official websites provided by the tRFMOs.

### The CWP Data Exchange Format

For effective amalgamation of multiple datasets, a unified data exchange format has been crafted. Central to this format is the inclusion of:

-   Digital Format Standardization:

The Comma-Separated-Value (CSV) format, adhering to the RFC 4180 standard, has been endorsed.

-   File Encoding Protocol:

Whether in CSV or JSON format, encoding is standardized to UTF-8.

This section describes the steps and procedures used to create the global nominal catch dataset from the data available within each tRFMO.

## Data call

Fisheries data sets managed by tRFMOs are generally updated on an annual basis according to their respective data submission cycle. Some revisions may also be performed on an irregular basis following corrections and improvements made to historical data. Each year, an official data call is made by the Steering Committee of the FAO Fisheries Resources Monitoring System (FIRMS) to get the most up-to-date versions of the nominal catch data set built by each Secretariat. It takes place around June to ensure that the data will include catches up to the year preceding the call.

<!-- In addition, each dataset can be periodically updated from the official websites of the tRFMOs. -->
<!-- Six datasets and five mapping datasets are used to create the final nominal catch dataset. -->

## Exchange format & mapping

- CWP standards
- Building of the specific GTA FIRMS fishing fleet code list 
- Mapping with ASFIS/ISSCFG/FIRMS fleets

<!-- ## Data pre-processing -->

<!-- Before combining the datasets, a pre-processing step is applied to three out of the five datasets to restructure their shape. This restructuring ensures uniformity in the format of the data, facilitating the subsequent integration process. Additionally, to ensure consistency and comparability across datasets, a harmonization process is performed to standardize the code lists. The harmonized datasets allow for seamless data integration and accurate analysis across different tRFMOs. -->

<!-- Below is a table listing the links to the datasets provided by each tRFMO: -->

<!-- ```{r, warning == FALSE, message = FALSE} -->
<!-- datasets_links_processing <- read_delim(here("inputs/Workflow/datasets_links_processing.csv"),  -->
<!--     delim = ";", escape_double = FALSE, col_types = cols(`Dataset;RFMO;Link to downolad;Pre-processing function;Comments` = col_character()), trim_ws = TRUE) -->
<!-- ``` -->

<!-- ```{r tab.cap = "Datasets and pre-processing functions used in the creation of the final dataset"} -->
<!-- qflextable(datasets_links_processing) -->
<!-- ``` -->

## Creation of the global dataset

Upon completing the pre-processing and harmonization steps, the global nominal catch dataset is constructed by binding and filtering the data.

<!-- ### Binding -->

<!-- All individual tRFMO datasets, representing different ocean regions, are combined through a data-binding process. This integration ensures comprehensive coverage of global tuna catch data, providing a holistic view of tuna fisheries on a global scale. The binding process brings together data from various regions, enabling cross-regional analysis and facilitating the identification of global trends and patterns in tuna catch. -->

### Filtering for southern Bluefin tuna (SBF) data from CCSBT

During this stage, a specific data filtering process is implemented to retain only the data relevant to Southern Bluefin Tuna (SBT) provided by the CCSBT (Commission for the Conservation of Southern Bluefin Tuna). As CCSBT possesses expertise in managing and overseeing data related to this species, their dataset is selected as the primary source for Southern Bluefin Tuna information. This careful selection ensures that the dataset remains focused and representative of SBT data while benefiting from the expertise of the CCSBT.

-   Data Dictionary:

A robust data dictionary has been established to guide the structure and semantics of the fisheries data exchange format. This repository, containing terms for column headers and their definitive explanations, is accessible at FDI Terms on GitHub. <https://github.com/fdiwg/fdi-terms> Notably, a specific fisheries term isn't rigidly bound to a single reference code list but varies based on its application context.

In context-specific scenarios, like the Global Tuna Atlas, associations with particular code lists : <https://github.com/fdiwg/fdi-mappings/tree/main/global/firms/gta>

Lastly, while two data structure definitions are being created, the Global Tuna Atlas has adopted the Generic (measurement-independent) data structure for its operations. \ref table

```{r results='asis'}
CWP_format <- read_delim("inputs/CWP_format.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
qflextable(CWP_format)
```

# Creation of the Global Dataset

## Harmonization

For now the CWP Data Exchange Format has not been adopted by all the tRFMOs yet. Then a pre processing part of Digital Format Standardization is ran before the workflow. These step is to deaseppear in future release. reftable\

```{r  results='asis'}
require(readr)
datasets_links_processing <- read_delim(here("inputs/Workflow/datasets_links_processing.csv"), 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
require(flextable)
qflextable(datasets_links_processing)
```

## Final output

Upon completing the CWP standardization, the global Nominal catch dataset is constructed in R. It takes in input the 5 datasets provided by tRFMOS and process 3 treatments.

1.  Binding of Datasets for each Ocean Region:

All the individual tRFMO datasets, representing different ocean regions, are combined through a data-binding process.

2.  Mapping the data

If the data is not mapped yet, the mapping is done on the basis of the codelist presented sooner. The recap of the mapping can be found in table \ref table

3.  Filtering on Species under Mandate by the tRFMOs:

This step involves filtering the data to retain only the information pertaining to species under the mandate of the tRFMOs (Regional Fisheries Management Organizations). The retained species are selected to align with the specific conservation and management goals of each tRFMO.

```{r}

file_url <- "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/cross-term/codelist_mapping_source_authority_species.csv"
download_link <- paste0("[Download CSV](<", file_url, ">)")

cl_species_level0 <- readr::read_csv(file_url)


```

```{r}

diffstratasnumber <- read.csv("outputs/filtering_species/figures/Diffstratas/Comparison.of.number.of.stratas.between.the.two.datasets.csv")

tidy <- diffstratasnumber[c(1, 3, 4, 5, 6),] %>% rename(" " = X, "Before filtering" = "Before.filtering", "After filtering" = "After.filtering")

qflextable2(tidy, columns_to_color = "Difference")

```

### Filtering on species under Mandate by the tRFMOs

```{r hide = TRUE, eval = FALSE}
file_url <- "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/cross-term/codelist_mapping_source_authority_species.csv"
download_link <- paste0("[Download CSV](<", file_url, ">)")

cl_species_level0 <- readr::read_csv(file_url)
```

This step involves filtering the data to retain only the information pertaining to species under the mandate of the tRFMOs. The retained species are carefully selected to align with the specific conservation and management goals of each tRFMO. The resulting dataset exclusively focuses on these species, ensuring a more targeted analysis that directly addresses the priorities and responsibilities of each respective tRFMO. `r SPECIES_ITIS[Aggregate == FALSE, length(unique(TSN))]` are selected. 

```{r tab.cap = "Species retained in the final dataset by tRFMOs", eval = FALSE}
require(dplyr)

ft <- flextable(as_grouped_data(cl_species_level0 %>% arrange(desc(species)) %>% dplyr::select(species, source_authority), groups = "species")) %>% 
  merge_v(j = "species", part = "header") %>%
  set_caption(caption = paste("Download Link:", download_link))

ft
```

## Data enrichment

To enhance the comprehensibility and usability of the dataset, additional temporal and grouping information is enriched in the data. Specifically, temporal information such as time stamps and date columns are added, enabling temporal analysis and trend identification. Additionally, group columns are introduced, categorizing species and fishing gear types, facilitating efficient grouping and analysis of the data based on common characteristics.

<!-- By enriching the dataset with these additional features, the resulting dataset becomes a valuable resource for researchers and policymakers to gain deeper insights into global tuna catch dynamics, fishing patterns, and species-specific trends. -->

<!-- The methods presented here enable the creation of a reliable, harmonized, and enriched Nominal catch dataset that contributes to a better understanding of global tuna fisheries.  -->


<!-- To be noted, one of the main tuna i.e. Southern Bluefin Tuna, for which CCSBT is considered the only authoritative source of information for the species is impacted by 4.3% of the value by this treatment. -->

To enhance the comprehensibility and usability of the dataset, additional temporal and grouping information is enriched in the data, categorizing species and fishing gear types using mapping of the Data Dictionary.

The code used to perform this workflow is available on: [https://raw.githubusercontent.com/eblondel/geoflow-tunaatlas/master/tunaatlas_scripts/generation/create_global_tuna_atlas_nominal_catch.R](#0){.uri}

## Geoflow workflow

The workflow of creation and the dissemination of the Nominal catch is using the geoflow package <https://github.com/r-geoflow/geoflow>.
