# Data Records

_The Data Records section should be used to explain each data record associated with this work, including the repository where this information is stored, and to provide an overview of the data files and their formats. Each external data record should be cited as described below. A data citation should also be placed in the subsection of the Methods containing the data-collection or procedure(s) used to derive the corresponding record._

_Tables can be used to support the data records, specifying the data output resulting from each data-collection or analytical step and the names of the corresponding files should these form part of the archived record._

## Data Records

This study encompasses several data records detailing global nominal catch figures across various regions and timelines. Each record represents the aggregate catch, specified by fish species, geographic_identifier, gear_type, fishing_fleet, and respective time periods.


```{r, warning == FALSE, message = FALSE}
datasets_links_processing <- read_delim(here("inputs/Workflow/datasets_links_processing.csv"),
    delim = ";", escape_double = FALSE, col_types = cols(`Dataset;RFMO;Link to downolad;Pre-processing function;Comments` = col_character()), trim_ws = TRUE)
```

```{r tab.cap = "Datasets and pre-processing functions used in the creation of the final dataset"}
qflextable(datasets_links_processing)
```



Data Repository
All data associated with this study are deposited and made publicly accessible in the , a googledrive for fisheries data . The dataset is persistently available at DOI.

The datasets used to create the data are available on a googledrive 

```{r}
if(!(file.exists(here("inputs/Workflow/nominal_files_recap.csv")))){
require(googledrive)
drive_auth()
require(googlesheets4)
nominal_files_recap <- read_sheet("https://docs.google.com/spreadsheets/d/13SJB2PYkCCdrzw8HNJMPFg0yXs0OSYY2mUOAl5dagUc/edit#gid=374572086")
fwrite(nominal_files_recap,here("inputs/Workflow/nominal_files_recap.csv" ))
} else {
  nominal_files_recap <- read_csv(here("inputs/Workflow/nominal_files_recap.csv"))

}
```


```{r tab.cap = "Recap of the files and their associated process used for the creation of the final output 'global_nominal_catch_firms_level0'"}
qflextable(nominal_files_recap)
```

This sheet is the input of the workflow through a json processed by the geoflow package. The first 3 lines concerned dataset being processed after receiving. The following two lines concerned data having minor processes has the format is already CWP. The last lines uses the 5 precedents outputs to create the final dataset. 

The codelist datasets make the link between mappings dataset and dimension of the initial dataset. 

### Dataset Overview

#### Pre processed data

##### Format

For the inputs data corresponding to the first 5 lines, all the formats are recapped in the code of the processed associated. The outputs format are always the CWP format.

##### Storage

The storage of the inputed files is done on the googledrive of firmstunaatlas, the output are stored in a database and in the googledrive. 

#### Final data

##### Format

The final dataset named : contains `r ifelse((length(colnames(NC)) == 24), length(colnames(NC)), FALSE)` columns. 11 of them corresponds to the CWP standards columns, the remaining 13 drift from an enrichment on the basis of the columns fishing_fleet, species, gear_type, geographic_identifier.


##### Storage

This data is stocked on zenodo under the DOI:
The geoflow package allows as well to store it in a googledrive, a database, a geoserver and a geonetwork catalogue.

##### Other files used for the display of maps, enrichment or filtering.

###### Filtering files

```{r}

file_url <- "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/cross-term/codelist_mapping_source_authority_species.csv"
download_link <- paste0("[Download CSV](<", file_url, ">)")

cl_species_level0 <- readr::read_csv(file_url)


```


```{r tab.cap = "Species retained in the final dataset by tRFMOs", eval = FALSE}
require(dplyr)

ft <- flextable(as_grouped_data(cl_species_level0 %>% arrange(desc(species)) %>% dplyr::select(species, source_authority), groups = "species")) %>% 
  merge_v(j = "species", part = "header") %>%
  set_caption(caption = paste("Download Link:", download_link))

ft
```


###### Enrichement files

```{r}

```

###### Shapes files
